---
title: "Class11"
author: "Juliette Bokor (PID: A16808121)"
format: pdf
---

## Section 1. Proportion of G/G in MXL population 

Downloaded a CSV file from Ensemble 
```{r}
mxl <- read.csv("MXL_populationSNP.csv")
head(mxl)
```


```{r}
table(mxl$Genotype..forward.strand.)
```
There are 9 samples with the genotype G|G





## Section 4: Homework on Population Scale Analysis

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
```{r}
nrow(expr) 
table(expr$geno)
```
```{r}
## I need to find the median for each genotype 
library(dplyr)

medians <- expr %>% 
  group_by(geno)%>%
  summarize(median_exp = median(exp))
medians
```


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2) 
exprplot <- ggplot(expr) + 
  aes(x=geno, y=exp, fill=geno) + 
  geom_boxplot(notch=TRUE)

exprplot
```


The relative expression of ORMDL3 is decreased with the G/G genotype in comparison to the 
A/A genotype. The SNP does affect the expression of ORMDL3. 

